% Please do not change the document class
\documentclass{scrartcl}

% Please do not change these packages
\usepackage[hidelinks]{hyperref}
\usepackage[none]{hyphenat}
\usepackage{setspace}
\doublespace

% You may add additional packages here
\usepackage{amsmath}

\title{How can different personality types best adapt to the agile philosophy?}

\subtitle{COMP150 - Agile Development Practice}

\author{1608351}

\begin{document}

\maketitle

\abstract{A variety of personalities and skill-sets in individuals within a team tends to enhance performance \cite{Paper3}, therefore it is important not to disregard those who struggle with agile principles. Finding methods that give individuals the autonomy to align their working style with agile, may help ensure diversity within teams.  This paper explores how individuals, with 
personality types that have been found to have a negative relationship with agile preference \cite{Paper1}, can best adapt to the agile philosophy.}

\section{Introduction}

Agile processes are widely used in the Game Development industry, and encourage a focus on self-organising teams along with regular reflection, and communication. Due to the scope involved in game development tasks, working within teams, rather than individually is necessary. However, such focus on teamwork and communication can be problematic for certain individuals. Researchers, D. Bishop, and A. Deokar, \cite{Paper1}, have found a correlation between an individual's personality type (based on the Five-Factor model) and their preference for agile processes, \textit{`The data supports a positive relationship between extraversion and agile preference as well as openness and agile preference'}. In contrast, subjects with introverted or neurotic personality types were less likely to feel inclined towards agile. Such findings are not unique to the Agile Development process, workplaces in general tend to favour those who are extroverted and open. As Susan Cain explains \cite{Paper6},\textit{`studies show that we rank fast and frequent talkers as more competent, likeable, and even smarter than slow ones' }. Such views are also seen in educational settings \textit{`research suggests that the majority of our teachers believe that the 'ideal student' is an extrovert.'} \cite{Paper6}.  This paper sets out to explore how introverted and neurotic personality types can best adapt to Agile and how they can be aided in doing so.
 
Whilst it could be argued that individuals who struggle with teamwork in general, are not suited to the game development industry, researchers such as L. Hoffman [] have found that a multiplicity of personalities are significantly beneficial to a team,\textit{ `The degree of homogeneity of personality of members of the groups used in this study was seen to have a direct bearing on the effectiveness of the groups in producing solutions to problems'}.  A diverse range of personality types can enhance team performance, so it is not only in the individual’s best interest but the team's.

\section{Introversion, Neuroticism and Agile}

Simply expecting an introverted or neurotic individual to adjust to agile is unlikely to be effective. Both personality types have traits that contradict with the agile manifesto. Introverted individuals tend to enjoy working with few interruptions, prefer not to show, or discuss work before it is completed and feel drained after only small amounts of social activity \cite{Paper6}. Whilst neurotic individuals tend to have higher levels of anxiety, fear, frustration and jealously.  These traits can make working within a team difficult, and within Agile Game Development, learning to manage such traits and preferences is essential.

Perhaps, it is important to remember that in adopting the Agile Manifesto, a team does not need to follow it as verbatim. Instead Agile can be adapted to teams by learning from both positive and negative experiences. 

\section{Potential Challenges}

To adapt to Agile effectively, individuals may benefit from identifying the agile processes they struggle with, and how particular traits influence these. For the purposes of this paper, I will be focusing on five key areas in which an introverted or neurotic individual may be reasonably expected to struggle. 
\newline
\newline
1.	\textbf{Accountability} \newline
In Agile, accountability ensures teams members are realistic with their workloads, contribute to the project, and helps increase trust within teams.\textit{`The accountability and collective responsibility in agile methods nurture trust by facilitating vigilance, aligning members' perceptions realistically with individual competences and abilities, fostering a sense of benevolence through team solidarity, and enhancing perceived integrity through a demonstrated shared work ethic'} \cite{Paper2}. However, both introverted and neurotic individuals may feel more pressure and anxiety in terms of accountability than their more open and extroverted colleagues. As observed by O. McHugh, K. Conboy and M. Lang in their study on trust in Agile teams, \textit{`“This pressure was entirely self-inflicted, but it might be consequent to the increased visibility of tasks and personal accountability to the team.'} \cite{Paper2} Thus, such team members may cope with their feelings of pressure through overworking in, or alternatively withdrawing from, the project.

2.	\textbf{Trust} \newline
Further to accountability, individuals with neurotic traits in particular, may find it difficult to form trust.

3.	\textbf{Social Demands} \newline
A key characteristic of introverts, is their need to have adequate time alone, often feeling drained by social activities. However, in the team focused enviroment of Agile Game Development, this can be problematic. As explained by E. Whitworth and R. Biddle, full immersion in teamwork and teambuilding activities helps to increase trust and a feeling of togetherness \textit{`Constant immersion and engagement with the team as a whole, for example, and the development of rituals surrounding team activity, were seen to support the development and prevalence of a shared identity.'} \cite{Paper7}, introverted individuals need to find a balance between the alone time they need and their interactions within the team.

4.	\textbf{Sharing} \newline

5.	\textbf{Communicationtion} \newline


\section{Conclusion}

Write your conclusion here. The conclusion should do more than summarise the essay, making clear the contribution of the work and highlighting key points, limitations, and outstanding questions. It should not introduce any new content or information.

\bibliographystyle{ieeetran}
\bibliography{agileessaybibliography}

\end{document}